% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[
]{article}
\usepackage{xcolor}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\newcounter{none} % for unnumbered tables
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}

\begin{document}

{\textbf{Welcome to Physics 12}}

{\textbf{}}\strut \\

{\textbf{}}\strut \\

{\textbf{}}\strut \\

{\textbf{}}\strut \\

{\textbf{Course Description:}}

This is an introductory course on physics that further explores the
basics of motion, forces, energy, fields, electricity, and optics.
Students will continue to build a foundation on scientific skills
necessarily for the study of physics. The course then focuses on the
topics above with applications to real life.{~ }The course is designed
to prepare students for university physics.

{\textbf{You Belong in Physics!}}

Physics might seem intimidating at first, but here\textquotesingle s the
truth: anyone can succeed in physics with curiosity, effort, and the
right support. This course is designed to build your confidence.

You don\textquotesingle t need to be a "math genius" to excel here.
Physics is about understanding how the world works, and
I\textquotesingle m here to help you develop both the problem-solving
skills and the confidence to tackle any challenge.

\textbf{Remember}: Every physicist started exactly where you are now.
Questions aren\textquotesingle t signs of weakness;
they\textquotesingle re signs of learning. I\textquotesingle m excited
to explore the fascinating world of physics with you this year!

\hfill\break

{\textbf{Contact Information:}}

Instructor:{~ }Mr. Gullo{~}

Option 1: Check your notes and textbook

Option 2: Ask a study partner or classmate

Option 3: Schoology messaging

Option 4: Office hours

{ }

{\textbf{Schoology:}}{ }

This powerful portal is where you will find all the course materials:{~
}PPTs, notes, review packages, etc.. as well as your grades

{\textbf{Course Materials}}

A textbook will be issued as a reference.{~}

Tests and quizzes are to be done on paper or schoology only.

Calculator: Casio fx-991EX ClassWiz (recommended)

A formula sheet will be given at the beginning of the year.{~ }This can
be used for tests, but no additional formulas added to the sheet.

\hfill\break

\textbf{Grading}

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}
  >{\raggedright\arraybackslash}p{(\linewidth - 4\tabcolsep) * \real{0.3333}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
\begin{minipage}[t]{\linewidth}\raggedright
\textbf{Unit Tests: 30\%}

\hfill\break

1 per unit

Each test 2 periods (80 min){~}\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
\textbf{Quizzes and Period Tests: 15\%}

\hfill\break

1-2 quizzes per unit{~ }(\textless{} 40 min)

1-2 period test per unit{~ }(40 min)

\hfill\break
\strut
\end{minipage} & \begin{minipage}[t]{\linewidth}\raggedright
\textbf{Activities/Labs: 15\%}

\hfill\break

1-2 per unit

1 lab per unit\strut
\end{minipage} \\
\begin{minipage}[t]{\linewidth}\raggedright
\textbf{Homework: 10\%}

{}\strut \\

1 check per unit\strut
\end{minipage} & \textbf{Midterm: 10\%} & \textbf{Final: 20\%} \\
\end{longtable}
}

{\textbf{}}\strut \\

{\textbf{Assessment for Learning and Written Work}}

-The \textbf{Assessment for Learning} category is worth \textbf{0\% of
your grade}. It is for practice and feedback only.

-On \textbf{quizzes}, you must show your work. If your steps are
correct, you can get partial marks even if the final answer is wrong.

On \textbf{unit tests,} \textbf{midterm and final exam}, written work
will be worth \textbf{30\% of the score}. Up to three questions
specified will be checked for written work.

\hfill\break

{\textbf{What We Will Learn: Topics Covered:}}

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}ll@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Unit & Chapters \\
Unit 1: Introduction to Physics 12 and Advanced Kinematics & 1,2,3 \\
Unit 2: Advanced Dynamics & 4,5,9 \\
Unit 3: Advanced Work, energy and power & 7 \\
Unit 4: Momentum and Conservation Laws & 8 \\
Unit 5: Circular Motion and Gravitation & 6 \\
Unit 6: Electrostatics and Electric Fields & 18,19 \\
Unit 7: Electric Circuits and Electromotive Force & 20,21 \\
Unit 8: Magnetism and Electromagnetic Induction & 22,23 \\
Unit 9: Special Relativity & 28 \\
\end{longtable}
}

\hfill\break

{~~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~
~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~ ~}

\hfill\break

\hfill\break

\hfill\break

{\textbf{}}\strut \\

{\textbf{}}\strut \\

{\textbf{Class Rules and Expectations:}}

{\textbf{Respect and Attitude:}}{~ }Treat the teacher, your classmates,
and most importantly yourself with respect. Believe in yourself and all
of your potential and capabilities. You have what it takes to succeed!{~
}Come to class with a positive attitude and do your best. The general
aim of this course is to build a community of young scientists who are
ready and willing to learn as well as help each other succeed.

{\textbf{Late/Attendance:}}{~ }Students are expected to attend all
classes, be punctual and come prepared. If a student is going to be
absent (drivers license, visa, etc.), please notify the teacher in
advance to help with scheduling. Unexcused absences lead to frustrations
when it comes to testing and this is unfair to other students. Multiple
absences will result in a talk to the homeroom teacher.

{\textbf{Deadlines:}} \textbf{} All assignments, labs, projects, etc.
are due at the {start} of the class on which they are due.{~}

{\textbf{Washrooms:}}{~ }It is expected that you will use the washroom
before or after class.{~ }If you need to use it during class, then do so
in a way that does not disrupt the lesson.

\textbf{Making Up Missed Assessments:}

I understand that sometimes life gets in the way of school.
Here\textquotesingle s how we\textquotesingle ll handle missed tests and
quizzes:

\textbf{Quizzes and Period Tests:} Makeup opportunities available with
advance notice when possible. See me to schedule.

\textbf{Unit Tests:} These are scheduled with at least one
week\textquotesingle s notice, so please plan accordingly. If multiple
students have conflicts, we can discuss rescheduling the test for
everyone.

\textbf{What to do if you miss an assessment:}

Talk to me as soon as you return

We\textquotesingle ll work together to find a solution

Bring any documentation if the absence was medical or family-related

{~}

\hfill\break

{\textbf{}}\strut \\

\textbf{Lab Safety}

Your safety is our top priority in all lab activities. All students must
follow these essential safety guidelines:

\textbf{Before Each Lab:}

Read all instructions carefully before beginning

Understand the procedure and ask questions if anything is unclear

Ensure long hair is tied back and loose clothing is secured

Remove or secure jewelry that could interfere with equipment

\textbf{During Lab Activities:}

Follow all written and verbal instructions

Never work alone - always work with your lab partner

Report any accidents, spills, or broken equipment immediately

Keep your workspace clean and organized

Handle all equipment with care

\textbf{Emergency Procedures:}

Know the location of safety equipment (fire extinguisher, first aid kit,
emergency exits)

Report any injury, no matter how minor, to the teacher immediately

In case of emergency, remain calm and follow teacher instructions

Emergency contact: School office at {[}extension number{]}

\textbf{Consequences:} Students who fail to follow safety procedures
will be removed from the lab activity and may receive a zero for that
assignment. Repeated safety violations will result in a meeting with
administration.

{\textbf{Cheating and Generative AI use:}} {~}

The Nanmo BC Academic Honesty Policy (found under our Course Admin
folder on Schoology) serves as our primary academic integrity framework
and supersedes all other policies. In addition to this policy, the
following specific procedures and guidelines will be implemented:

1. In cases of suspected cheating, while we maintain the principle of
\textquotesingle innocent until proven guilty\textquotesingle{} students
may be asked to demonstrate their understanding by answering similar
questions to verify their knowledge. Inability to demonstrate
understanding will result in a mark of zero.

2. Assisting others in cheating will also result in a mark of zero.

3. Repeat offenses will be referred to the principal.

4. Regarding Generative AI (like ChatGPT and others): While these are
valuable academic tools for concept explanation and understanding, they
can also provide inaccurate information. {\textbf{Submitting
AI-generated work as your own violates our Academic Honesty Policy.}}
Use these tools cautiously and responsibly.

\hfill\break

{\textbf{Sleeping}}{~ }It is considered a good habit to get sufficient
sleep every night.{~ }Part of that means limiting the use of devices and
avoiding video games.{~ }It also means good planning and time management
to deal with coursework timely and not stay up late.{~ }\textbf{Class
time is not a chance to make up lost sleep}.{~ }For those students who
repeatedly try and sleep in class, we will discuss this with the
homeroom teacher to find solutions.

{\textbf{Re-submission and Late-submission Policy}}{~}

1. Students are permitted to request a resubmission of one assessment
that received a grade under 60\% one time each semester via the
microsoft form {[}Deadlines two weeks before semester end{]}. The
highest possible grade achievable on a resubmitted assignment is capped
at 75\%, to maintain fairness and uphold academic standards, while still
providing a chance for improvement.

2. Late submissions will assigned as ``missing'' and graded zero unless
an extension is discussed with the instructor before the original due
date.

3. All student \textbf{submissions must be a single PDF or PPT file},
otherwise they will receive a grade of zero.

{\textbf{Messaging Procedures for Schoology}}

\textbf{Subject Line:} Physics 12 - {[}Your Section{]} - {[}Your
Block{]}{~}

Example: "Physics 12 - Section 2 - Block C"

\textbf{For homework or concept questions:}

Tell me which unit or assignment you\textquotesingle re working on

Show me what you\textquotesingle ve tried so far (attach a photo of your
work)

Ask your specific question

\textbf{For lab questions:}

Name the lab and your group number

Tell me what part you need help with

Include any data or calculations if relevant

\textbf{Response time:} I\textquotesingle ll get back to you within 24
hours on school days, next school day for weekend messages.

{\textbf{Electronic Device Use Policy}}{~ ~}

We will adhere to the Nanmo BC Electronic Device Use Policy found under
our Course Admin folder on Schoology.{~}

\hfill\break

{\textbf{Classroom Cell Phone Policy}}

\hfill\break

All cell phones must either be placed in the designated phone box or
stored in student lockers during class time. No phones are permitted to
remain with students during class unless explicitly approved by the
teacher.

\hfill\break

\textbf{Phone Box Procedure}

1. Upon entering class, students who have their phones with them must:

{~~ }- Place their phone in the designated phone box

{~~ }- Ensure their phone is powered off or on silent

{~~}

2. If a student does not place their phone in the box:

{~~ }- It is assumed their phone is stored in their locker

{~~ }- No phones may remain in bags, pockets, or desks

\hfill\break

\textbf{Phone Usage}

- Phones may only be retrieved from the box with explicit teacher
approval

- Students must state their reason for needing phone access

- Phones must be returned to the box immediately after approved use

- Emergency calls should be directed through the school office

\hfill\break

\textbf{Honor System Guidelines}

- Students are trusted to follow this policy honestly

- No individual phone checks will be conducted

- Students are expected to demonstrate integrity by:

{~ }- Either placing phones in the box or leaving them in lockers

{~ }- Not accessing phones without permission

{~ }- Being truthful about phone possession and usage

\hfill\break

\textbf{Consequences}

If a student is found using or possessing a phone during class without
approval:

1. The phone will be confiscated

2. School disciplinary policy regarding unauthorized phone use will be
followed

\hfill\break

Student Name:\_\_\_\_\_\_\_\_\_\_\_\_\_\_\_

\hfill\break

\hfill\break

\hfill\break

\end{document}
