% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[]{article}
\usepackage{xcolor}
\usepackage{graphicx}
\graphicspath{{docx-images/}{../images/}{../../shared/images/}}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\usepackage{longtable,booktabs,array}
\newcounter{none} % for unnumbered tables
\usepackage{calc} % for calculating minipage widths
% Correct order of tables after \paragraph or \subparagraph
\usepackage{etoolbox}
\makeatletter
\patchcmd\longtable{\par}{\if@noskipsec\mbox{}\fi\par}{}{}
\makeatother
% Allow footnotes in longtable head/foot
\IfFileExists{footnotehyper.sty}{\usepackage{footnotehyper}}{\usepackage{footnote}}
\makesavenoteenv{longtable}
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}

\begin{document}

Problem-Solving Rubric

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1118}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2508}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2120}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2125}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.2129}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Criteria & Extending & Proficient & Developing & Emerging \\
Problem Setup & Creates comprehensive, precise sketches with detailed
labels and annotations. Systematically identifies all relevant
quantities and relationships. Shows insight in defining system
boundaries. & Creates clear, appropriate sketches with necessary labels.
Correctly identifies known and unknown quantities. Appropriately defines
system of interest. & Creates basic sketches with some labels.
Identifies some relevant quantities. Shows partial understanding of
system boundaries. & Creates minimal or unclear sketches. Shows limited
recognition of relevant quantities. Unclear understanding of system
boundaries. \\
Free-Body Diagram & Constructs precise diagrams with exemplary detail.
Shows sophisticated understanding through accurate vector
representation. Demonstrates thorough understanding of vector component
resolution. & Constructs accurate free-body diagrams. Represents forces
correctly as vectors. Resolves vectors into components appropriately. &
Constructs basic diagrams with some errors. Shows inconsistent
understanding of force representation. Attempts vector resolution with
partial success. & Attempts basic diagrams with significant errors.
Shows beginning understanding of force representation. Limited
understanding of vector components. \\
Mathematical Analysis & Applies Newton\textquotesingle s Second Law with
precision in all directions. Shows deep understanding of force
relationships. Demonstrates elegant and efficient solution strategies. &
Applies Newton\textquotesingle s Second Law correctly. Sets up equations
properly. Follows logical solution process. & Shows developing
understanding of Newton\textquotesingle s Second Law. Sets up equations
with some errors. Solution process shows gaps in understanding. & Shows
emerging understanding of Newton\textquotesingle s Second Law. Equation
setup needs significant support. Solution process needs substantial
guidance. \\
Solution Verification & Provides insightful analysis of answer
reasonableness. Shows meticulous attention to units and significant
figures. Connects solution to broader physics principles. & Verifies
reasonableness of solution. Maintains correct units throughout. Uses
appropriate significant figures. & Attempts to verify solutions with
limited success. Shows inconsistent use of units. Limited attention to
significant figures. & Limited awareness of solution verification.
Inconsistent use of units. Minimal attention to significant figures. \\
\end{longtable}
}

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.1292}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.5310}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.3280}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.0054}}
  >{\raggedright\arraybackslash}p{(\linewidth - 8\tabcolsep) * \real{0.0063}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Criteria & Extending & Proficient & & \\
Problem Setup & Creates comprehensive, precise sketches with detailed
labels and annotations.

Systematically identifies all relevant quantities and relationships.
Shows insight in defining system boundaries. & Creates clear,
appropriate sketches with necessary labels.

Correctly identifies known and unknown quantities. Appropriately defines
system of interest. & & \\
Free-Body Diagram & Constructs precise diagrams with exemplary detail.

Shows sophisticated understanding through accurate vector
representation.

Demonstrates thorough understanding of vector component resolution. &
Constructs accurate free-body diagrams.

Represents forces correctly as vectors.

Resolves vectors into components appropriately. & & \\
Mathematical Analysis & Applies Newton\textquotesingle s Second Law with
precision in all directions. Shows deep understanding of force
relationships. Demonstrates elegant and efficient solution strategies. &
Applies Newton\textquotesingle s Second Law correctly.

Sets up equations properly.

Follows logical solution process. & & \\
Solution Verification & Provides insightful analysis of answer
reasonableness.

Shows meticulous attention to units and significant figures. Connects
solution to broader physics principles. & Verifies reasonableness of
solution.

Maintains correct units throughout.

Uses appropriate significant figures. & & \\
\end{longtable}
}

{\def\LTcaptype{none} % do not increment counter
\begin{longtable}[]{@{}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.1212}}
  >{\raggedright\arraybackslash}p{(\linewidth - 2\tabcolsep) * \real{0.8788}}@{}}
\toprule\noalign{}
\endhead
\bottomrule\noalign{}
\endlastfoot
Criteria & Extending \\
Problem Setup & Creates comprehensive, precise sketches with detailed
labels and annotations.

Systematically identifies all relevant quantities and relationships.

Shows insight in defining system boundaries. \\
Free-Body Diagram & Constructs precise diagrams with exemplary detail.

Shows sophisticated understanding through accurate vector
representation.

Demonstrates thorough understanding of vector component resolution. \\
Mathematical Analysis & Applies Newton\textquotesingle s Second Law with
precision in all directions.

Shows deep understanding of force relationships.

Demonstrates elegant and efficient solution strategies. \\
Solution Verification & Provides insightful analysis of answer
reasonableness.

Shows meticulous attention to units and significant figures.

Connects solution to broader physics principles. \\
\end{longtable}
}

\end{document}
