\documentclass[12pt]{article}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath}
\usepackage{graphicx}
\graphicspath{{../images/}{../../shared/images/}}
\usepackage{enumitem}
\usepackage{fancyhdr}
\usepackage{tcolorbox}
\usepackage{hyperref}
\usepackage{longtable}
\usepackage{array}
\usepackage{booktabs}

\pagestyle{fancy}
\fancyhead{}
\fancyhead[L]{Physics 11}
\fancyhead[R]{FPPL \& Core Competencies}
\fancyfoot{}
\fancyfoot[C]{\thepage}

\begin{document}

\begin{center}
\Large\textbf{Physics Reflection: FPPL \& Core Competencies}\\[0.5em]
\normalsize Physics 11 - Mr. Gullo
\end{center}

\vspace{0.5em}
Names: \underline{\hspace{8cm}}

\section*{Activity 1: Physics + First Peoples Principles Mind Map}

\textbf{Complete on FRONT of A3 paper}

\subsection*{Step 1: Think Alone (10 min)}

Review your completed work. Find connections to First Peoples Principles.

\textbf{My FPPL Connections:}

\begin{center}
\begin{tabular}{|c|p{10cm}|}
\hline
\textbf{Unit/Project} & \textbf{FPPL Connection (land, community, observation, relationships)} \\
\hline
1. & \\[1cm]
\hline
2. & \\[1cm]
\hline
3. & \\[1cm]
\hline
\end{tabular}
\end{center}

\subsection*{Step 2: Pair Work (15 min)}

On your A3 paper FRONT side:

\begin{itemize}
\item Draw ``Physics Through FPPL'' in center
\item Add branches for each unit you studied
\item Write your FPPL connections from table above
\item Draw arrows showing relationships
\item Use whole page
\end{itemize}

\subsection*{Step 3: Gallery Walk (10 min)}

Walk around. Use sticky notes on other posters. What stood out?

\textbf{Feedback we received:}

\vspace{2cm}

\textbf{(TURN PAGE OVER)}

\newpage

\section*{Activity 2: Core Competencies Evidence Board}

\textbf{Complete on BACK of A3 paper}

\subsection*{Step 1: Think Alone (12 min)}

\textbf{My Evidence:}

\begin{center}
\begin{tabular}{|p{3cm}|p{4cm}|p{4cm}|c|}
\hline
\textbf{Physics Task} & \textbf{Competency Used (T/C/PS)} & \textbf{Proof (what I did)} & \textbf{Growth 1-5} \\
\hline
1. & & & \\[1cm]
\hline
2. & & & \\[1cm]
\hline
3. & & & \\[1cm]
\hline
\end{tabular}
\end{center}

\textbf{T} = Thinking, \textbf{C} = Communication, \textbf{PS} = Personal/Social

\subsection*{Step 2: Pair Work (15 min)}

On your A3 paper BACK side, divide into 3 sections:

\textbf{Thinking | Communication | Personal/Social}

In each section:
\begin{itemize}
\item Write examples from both partners (use initials)
\item Brief description of competency shown
\item Draw arrows showing growth over time
\end{itemize}

\subsection*{Step 3: Speed Share (8 min)}

Rotate to other pairs. Explain your board (2 min each pair).

\textbf{One insight from another pair's board:}

\vspace{3cm}

\end{document}
