% Options for packages loaded elsewhere
\PassOptionsToPackage{unicode}{hyperref}
\PassOptionsToPackage{hyphens}{url}
\documentclass[]{article}
\usepackage{xcolor}
\usepackage{graphicx}
\graphicspath{{docx-images/}{../images/}{../../shared/images/}}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb}
\setcounter{secnumdepth}{-\maxdimen} % remove section numbering
\usepackage{iftex}
\ifPDFTeX
  \usepackage[T1]{fontenc}
  \usepackage[utf8]{inputenc}
  \usepackage{textcomp} % provide euro and other symbols
\else % if luatex or xetex
  \usepackage{unicode-math} % this also loads fontspec
  \defaultfontfeatures{Scale=MatchLowercase}
  \defaultfontfeatures[\rmfamily]{Ligatures=TeX,Scale=1}
\fi
\usepackage{lmodern}
\ifPDFTeX\else
  % xetex/luatex font selection
\fi
% Use upquote if available, for straight quotes in verbatim environments
\IfFileExists{upquote.sty}{\usepackage{upquote}}{}
\IfFileExists{microtype.sty}{% use microtype if available
  \usepackage[]{microtype}
  \UseMicrotypeSet[protrusion]{basicmath} % disable protrusion for tt fonts
}{}
\makeatletter
\@ifundefined{KOMAClassName}{% if non-KOMA class
  \IfFileExists{parskip.sty}{%
    \usepackage{parskip}
  }{% else
    \setlength{\parindent}{0pt}
    \setlength{\parskip}{6pt plus 2pt minus 1pt}}
}{% if KOMA class
  \KOMAoptions{parskip=half}}
\makeatother
\setlength{\emergencystretch}{3em} % prevent overfull lines
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{bookmark}
\IfFileExists{xurl.sty}{\usepackage{xurl}}{} % add URL line breaks if available
\urlstyle{same}
\hypersetup{
  hidelinks,
  pdfcreator={LaTeX via pandoc}}

\author{}
\date{}

\begin{document}

\textbf{Lab Design Assignment: Gravitational Force Investigation with
Steel Ball Drop}

\textbf{Overview}

This experimental investigation has been structured to fit an 80-minute
lab period, with an additional 40-minute design period the week before.
Students will use a timer-controlled dropping mechanism and photo-gate
to study concepts from Chapter 6 on uniform circular motion and
gravitation, specifically focusing on determining the gravitational
force between the Earth and the steel ball.

\textbf{Schedule}

Week 1: Design Period (40 minutes)

Week 2: Lab Implementation (80 minutes)

\textbf{Required Equipment}

\begin{itemize}
\item
  Timer-controlled dropping mechanism
\item
  Steel ball
\item
  Photo-gate timing system
\item
  Electronic balance
\item
  Calculator
\item
  Laptop with spreadsheet software (optional)
\item
  Lab notebook
\end{itemize}

\textbf{\hfill\break
}

\textbf{Pre-Lab Design Requirements (Week 1)}

During the 40-minute design period, students should:

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Review textbook Chapter 6 to identify relevant equations, especially
  those related to gravitational force and acceleration due to gravity.
\item
  Design a systematic approach for data collection that will allow them
  to calculate the acceleration due to gravity.
\item
  Create a data table template that includes measured height, time, and
  calculated acceleration.
\item
  Develop a hypothesis about how the measured acceleration due to
  gravity will compare to the accepted value.
\item
  Prepare prediction calculations for the acceleration due to gravity
  based on the timing data at different drop heights.
\item
  Identify potential sources of experimental error that could affect the
  measured acceleration.
\end{enumerate}

\textbf{\hfill\break
}

\textbf{Experimental Procedure (Week 2)}

\textbf{Setup Phase (15 minutes)}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Measure and record the mass of the steel ball.
\item
  Mount the dropping mechanism securely above the photo-gate.
\item
  Test the timing system and dropping mechanism.
\item
  Establish a consistent method for measuring drop heights.
\end{enumerate}

\textbf{Data Collection Phase (30 minutes)}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Select 5 different drop heights between 0.2m and 1.0m.
\item
  For each height:

  \begin{itemize}
  \item
    Record the height measurement precisely.
  \item
    Perform 3 trial drops, recording the photo-gate time measurements.
  \item
    Calculate the acceleration due to gravity (g) using relevant
    equations and the measured values.
  \end{itemize}
\end{enumerate}

\textbf{Analysis Phase (20 minutes)}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Calculate the average acceleration due to gravity (g) for each drop
  height.
\item
  Calculate the gravitational force (Fg) acting on the ball using the
  measured mass and calculated acceleration (Fg = mg).
\item
  Calculate the percent error between the average measured acceleration
  and the accepted value of g (9.8 m/s²).
\item
  Create a graph of drop height vs. calculated acceleration due to
  gravity.
\item
  Analyze any trends observed in the data.
\end{enumerate}

\textbf{Discussion Phase (10 minutes)}

\begin{enumerate}
\def\labelenumi{\arabic{enumi}.}
\item
  Compare the calculated gravitational force to the expected force.
\item
  Analyze discrepancies between measured and accepted values of g.
\item
  Discuss experimental uncertainties and their impact on the calculated
  gravitational force.
\item
  Relate findings to Newton\textquotesingle s law of universal
  gravitation and the concept of acceleration due to gravity.
\end{enumerate}

\textbf{Submit:}

\begin{itemize}
\item
  \textbf{By end of design period (Week 1):}

  \begin{itemize}
  \item
    Completed experimental design
  \item
    Prediction calculations for force due to gravity.
  \end{itemize}
\item
  \textbf{By end of lab period (Week 2):}

  \begin{itemize}
  \item
    Completed data tables
  \item
    Preliminary graphs
  \item
    Initial analysis of results.
  \end{itemize}
\item
  \textbf{Final report (due one week after lab):}

  \begin{itemize}
  \item
    Introduction with hypothesis
  \item
    Methods section with detailed procedure
  \item
    Results with data tables and graphs
  \item
    Analysis connecting to Newton\textquotesingle s law of universal
    gravitation
  \item
    Discussion of error sources
  \item
    Conclusion evaluating the hypothesis
  \end{itemize}
\end{itemize}

\textbf{Safety Considerations}

\begin{itemize}
\item
  Secure all equipment properly to prevent falling objects.
\item
  Ensure clear space around the drop zone.
\item
  Handle equipment with care to prevent damage.
\end{itemize}

\section{Physics Laboratory Skills Proficiency
Rubric}\label{physics-laboratory-skills-proficiency-rubric}

\subsection{Emerging}\label{emerging}

Description: Students at this level require significant guidance to
complete laboratory tasks. They demonstrate basic understanding of
physics concepts but struggle to connect theory with experimental
practice. Work is primarily procedural and follows step-by-step
instructions with minimal independent thinking.

Skills and Abilities:

\begin{itemize}
\item
  Sets up basic laboratory equipment with teacher assistance and follows
  preset procedures
\item
  Records data with limited attention to precision or measurement
  techniques
\item
  Performs simple calculations using provided formulas without deep
  understanding of their significance
\item
  Identifies obvious sources of error but struggles to analyze their
  impact on results
\end{itemize}

\subsection{Developing}\label{developing}

Description: Students at this level can perform laboratory tasks with
some independence. They show a growing understanding of physics concepts
and can apply basic theory to experiments. They require occasional
guidance but can follow procedures and record data systematically.

Skills and Abilities:

\begin{itemize}
\item
  Sets up laboratory equipment correctly with minimal assistance and can
  calibrate basic instruments
\item
  Records data systematically with appropriate units and reasonable
  precision
\item
  Performs calculations correctly and begins to recognize the
  relationship between variables
\item
  Identifies several sources of experimental error and attempts basic
  quantitative error analysis
\end{itemize}

\subsection{Proficient}\label{proficient}

Description: Students at this level work independently in the laboratory
setting. They demonstrate solid understanding of physics concepts and
can apply theoretical principles to design and implement experiments.
They work methodically, maintain accuracy in their measurements, and can
analyze results in relation to physical laws.

Skills and Abilities:

\begin{itemize}
\item
  Sets up complex experimental apparatus independently and troubleshoots
  equipment issues
\item
  Collects data with appropriate precision, recognizing the limitations
  of measuring instruments
\item
  Performs comprehensive data analysis, including appropriate graphs and
  trend identification
\item
  Conducts thorough error analysis and can distinguish between
  systematic and random errors
\end{itemize}

\subsection{Extending}\label{extending}

Description: Students at this level demonstrate exceptional laboratory
skills and conceptual understanding. They work with complete
independence, often extending investigations beyond requirements. They
show sophisticated understanding of physical principles and can design
novel experimental approaches to test hypotheses or explore
relationships between variables.

Skills and Abilities:

\begin{itemize}
\item
  Designs and implements original experimental procedures to investigate
  physics concepts
\item
  Develops innovative data collection strategies to maximize precision
  and minimize uncertainty
\item
  Performs advanced analysis connecting multiple physics concepts to
  explain experimental results
\item
  Applies critical thinking to evaluate experimental design and proposes
  refinements based on theoretical principles
\end{itemize}

\end{document}
